\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-24 europa.tex
addressed pending items from our last conversation. a number of fixes after a final review.

Fig. \ref{fig:europa-architecture} shows the main architectural
components in \eu and their relationships. The \textbf{Constraint
  Reasoning Engine} (\texttt{CRE}) manages domain types, variables and
constraints that define relationships among them. It also provides an
efficient arc consistency mechanism \cite{mackworth77}. The
\texttt{CRE} is designed so that specialized reasoning algorithms for
new constraints can be easily and efficiently plugged in. The
\textbf{Plan Database} (\texttt{PDB}) manages object and token
types. It also maintains the state of partial plans in terms of
object, variable and token instances. This constitutes the data
foundation on which end user applications and automated planners
within \eu are built. The \textbf{Rules Engine} (\texttt{RE})
modularly adds the semantics of domain rules that describe
dependencies between tokens. The \textbf{Solver Module} provides the
framework for flaw detection and resolution that is needed to
implement search algorithms for planning. The \textbf{API Layer}
provides access to services from all the different modules in a
programmatic way so that \eu can be easily embedded into specialized
applications.

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eue,
\texttt{NDDL} allows the user to specify the different components of a
problem domain in a precise and concise manner. Among the main
features that make \texttt{NDDL} a powerful tool for describing
problem domains are that it is object oriented. \texttt{NDDL} supports
classes and inheritance in similar fashion to popular Object-Oriented
languages like Java and C++. It also supports polymorphism for some of
the planning components. Using object classes and instances is a
time-tested approach to naturally describe problem domains
\cite{rumbaugh}. It offers declarative constructs to define
constraints and causality dependencies between actions and effects and
finally it offers procedural constructs to populate \eus \texttt{PDB}
with specific problem instances expressed in terms of objects,
variables, constraints, facts and goals. To articulate important
elements of \texttt{NDDL}, we show how the Shopping Agent problem
could be modeled:

@@

In the model above one can see the main elements of a problem domain
can be mapped to \eus representation:

@@

\item[\textbf{Object Types}]: \texttt{NDDL} supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. \texttt{NDDL} supports a single rooted class
  hierarchy so that state and behavior can be extended through
  inheritance. We take advantage of that feature by making
  \textit{AgentLocation} inherit from \eus built-in \textit{Timeline}
  class which doesn't allow predicates to overlap in time for the same
  object instance. This feature is used for instance, to ensure that
  an agent can only be at one place at a time in our example.

\item[\textbf{Variable Types and Instances}]: All object attributes
  are represented as variables in this model. For instance, it uses
  integer variables to represent the built-in temporal parameters for
  each token (activity duration, start and end times), string
  variables to represent \textit{Location} and \textit{Product} names,
  and object reference variables to represent the structural
  relationships between different objects (for instance, each Agent
  maintains a variable to keep track of its \textit{AgentLocation}).

\item[\textbf{Predicate Types}]: Some classes like \textit{Product},
  \textit{Location} and \textit{ProductLocation} are static and so
  don't have any predicate or action types associated with them. The
  \textit{AgentLocation} class is used to keep track of state and so
  has \texttt{At} and \texttt{Going} predicate types associated with
  it. The \textit{Agent} class keeps track of the products it
  purchases through the \texttt{Own} predicate; having a separate
  \textit{AgentBag} class with its own predicates would have been an
  equally valid modeling choice.

\item[\textbf{Action Types}]: In our example, \textit{Agent} is solely
  the class that has behavior associated with it. Consequently, it
  defines action types to represent when the agent is taking action to
  \texttt{Go} to a \textit{Location}, or to \texttt{Buy} a
  \textit{Product}. % We'll see how the actual behavior for those
  % actions is specified below.

@@

\item[\textbf{Variables}]: A token type (predicate or action) is part
  of a class, therefore class attributes (see \texttt{object.location}
  above), token type parameters (see \texttt{[from, to]} in
  \texttt{Agent::Go}, or \texttt{[duration, product]} in
  \texttt{Agent::Buy}) and locally declared variables (see
  \texttt{possibleStores} in \texttt{Agent::Buy}) can all be used in
  the definition of a rule.

\item[\textbf{Constraints}]: Restrictions on values that variables can
  take within a rule are expressed through constraints. In the example
  above the \texttt{eq} constraint is used; in general, any constraint
  available in \eus library can be used.

\item[\textbf{Subgoals}]: In \texttt{Agent::Go}, subgoals are used to
  specify that the Shopping Agent's location must transition from
  origin to destination in a temporally consistent way. The
  \texttt{[meets, equals, metby]} operators (see Section
  \ref{sec:europa:inference}) are used to specify subgoals, along with
  the temporal relationship each token representing a subgoal has with
  the the token representing the \texttt{Agent::Go} action. Also,
  \texttt{[condition, effect]} annotations can be used when defining a
  subgoal inside an action. Currently \eu does not associate semantics
  with these annotation; but they can be extremely useful for search
  modules when creating plans.

@@

\label{eu:model}
\end{verbatim}

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts are used to specify the current state of the
world and do not need any support and are represented as tokens in the
\texttt{PDB}. In this case we use facts to specify that the Shipping
Agent is initially at home. Goals are also represented as tokens in
the \texttt{PDB}. Different from facts, goals trigger inference in \eu
which determines whether flaws (unresolved subgoals, overloaded
resources, etc.) need to be resolved; if so, search modules can be
invoked to resolve those flaws as part of the deliberation
process. Finally, constraints are restrictions on and between actions
and goals. In the example above, we use temporal constraints to
specify when facts hold and when the goals should be
accomplished. % As
% always, all constraints in \eus library could be used.

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} relations shown above.}
\label{fig:allen-algebra}
\end{figure}

In \eu, problem domains use \texttt{Allen Algebra} \cite{allen84}
temporal relations shown in Fig. \ref{fig:allen-algebra} which allow
encoding of temporal constraint relationships between tokens. 
%\kcomment{This para could do with a little
  embellishment especially since it is the only material that actually
  talks about the kinds of constraints and their origin.JB: I don't really know what else to say given the table above.}

@@

In \eue, all variables and constraints from a problem domain
description are stored inside the \texttt{CRE}.  The \texttt{CRE}
provides a framework where any number of constraint types can
collaborate to perform inference and prune variable domains using
Bounds Propagation and Domain Reduction techniques \cite{marriott98}. 
\eu uses a Propagator pattern which keeps track of a set of
constraint instances of the same type and is responsible for ensuring
variables are inferentially updated upon change (domain modifications or
constraint addition/deletion).

% that whenever any variable is modified or constraints are introduced
% or dropped, all related variables are updated through inference.

In a planning problem, variable or constraint changes can come
directly from an end user (by introducing new activities or goals and
relating them to the rest of the plan through constraints, by changing
the start time of an activity, etc), or from an automated planner (by
expanding subgoals, introducing precedence constraints to resolve
mutual exclusion conflicts, etc). When these changes occur, each
propagator is only able to ensure consistency for the constraints that
are associated with it.  To maintain global consistency, the
\texttt{CRE} implements a variant of the \texttt{AC-3} algorithm
\cite{mackworth77} where propagators are triggered to infer to ensure
the constraint network returns to a quiescent state.

@@

An \eu user can introduce new constraint types by implementing
specialized constraint and propagator classes; however, \eu provides a
built-in constraint library with a range of arithmetic, temporal, set
and resource constraint types so that a large number of problem
domains can be modeled without having to introduce new constraint
types. Tables \ref{tab:calconst}, \ref{tab:setconst},
\ref{tab:objhierarchy} and \ref{tab:misc} show various constraint
types.

@@

Temporal and resource constraints are especially important to model
real-world planning domains. % so it is worth going into some more
% detail about how \eus built-in support works for those constraint
% types.
\eu uses variables to explicitly represent timepoints for plan
activities and states. Constraints among timepoints provide a natural
way to express domain axioms. For example, in order to state that
activity $A$ must occur before activity $B$ we can say that the end
timepoint of $A$ is $leq$ the start timepoint of $B$.
\cite{dechter91} proposed that constraints among timepoints can be
grouped together to form a Simple Temporal Network
(\texttt{STN}). Such a network can be transformed into a Distance
Graph (\texttt{DG}) where the outward arc from a node represents the
maximum distance from the source node to the target node.
Fig. \ref{fig:stn} illustrates an \texttt{STN} with $2$ variables and
a single constraint along with the resulting \texttt{DG}.
\cite{dechter91} also showed that shortest path algorithms could be
used to propagate values in the network and discover a negative cycle:
namely a path from a node to itself that has a path length less than
$0$. If such a cycle exists, the network is inconsistent. It was
further shown that a single-source shortest path algorithm was
sufficient to detect a negative cycle and provide sufficient
propagation to yield a backtrack-free search. \eus propagator for
temporal constraints is a straight implementation of this algorithm.

@@

Most real-world planning problems need to deal with plans made up of
actions that typically depend upon limited resources (\eg fuel,
storage capacity, battery state of charge, data buffers).  \eu
provides built-in support for resource computation. In \eus
representation resource object types come in three forms with related
constraints. \textbf{Reservoir} resources can have both
\texttt{consume(resource,time,quantity)} and
\texttt{produce(resource,time,quantity)} constraints; production and
consumption is instantaneous. \textbf{Reusable} resources on the other
hand, have constraints that use a quantity of the resource for a
specific duration (\texttt{uses(resource,start,duration,quantity)}),
which are then subsequently released. \textbf {Unary} resources are
reusable with capacity$ = 1$, which allows \eu to perform run time
savings in comparison to the generic reusable class. Configuration
options are offered such that in addition to the usual constraint
violations when a resource is completely depleted, plan flaws can be
generated whenever a resource level is outside a specific upper or
lower bound, or when instantaneous production or consumption exceeds a
specific threshold.

Since \eu supports a flexible temporal representation, where time is
represented by intervals, instead of single values, determining
whether a resource is the source of a plan flaw or a constraint
violation in general is substantially more difficult than just
computing a profile of the resource and checking against some limits.
With temporal flexibility, a number of resource profiles could be
possible depending on precisely when each production/consumption event
occurs out of a range of possible values that are in the domain of the
variables that represent the temporal extant for the event.
\cite{Muscettola04,Muscettola06} proposed an algorithm where the
variables that represent the time for production/consumption events
are placed in a temporal network, similar to the network used by the
temporal propagator; constraints on the resource can then be seen as
quantities flowing through that network. Computing the maximum flow
then, yields the maximum resource usage and thus a lower bound on the
resource level; computing a minimum flow will similarly yield an upper
bound. Since maximum flow on a temporal network can be computed in
polynomial time, \eu takes advantage of this algorithm to efficiently
compute upper and lower bound envelopes on the resource
level. Fig. \ref{fig:resenvelopes} shows an example of such a resource
computation.

@@

Resource envelopes can be useful when trying to evaluate plan
brittleness in the face of changing activity durations or start
times. They can also be useful when searching for plans; however they
are relatively expensive to compute. Therefore their ability to prune
the search space must be weighted against the ability to search faster
\cite{6007772}.

@@

Given the above, an automated problem solver can be created using, a
\emph{domain model} that describes the variable, object, predicate and
action types that are relevant for the problem, a \emph{problem
  instance} or initial state that consists of variable and object
instances that exist for the entire planning horizon, \emph{temporally
  scoped predicate and action} instances and \emph{temporally scoped
  goals}. This information is retained in \eus \texttt{PDB} so that
inference and search mechanisms can be used to look for a problem
solution.

Since temporal intervals may extend indefinitely in both positive and
negative directions, for problems with a temporal dimension it is also
common to specify a \emph{planning horizon} when invoking a solver;
any decisions that fall completely outside of the horizon can be
ignored. This is particularly useful for recurrent activities;
consider for instance a model where a medical checkup has to occur
every $3$ months. The user can model a recurring activity and then use
the horizon to ensure that a finite number of medical checkup
activities are generated as part of a plan. This is also useful to
ignore activities that may occur far in the future or in the past to
be relevant for a particular plan, ensuring that the solver is making
as few commitments as necessary and making planning tractable.

\subsubsection{Flaw Resolution}
\label{sec:europa:flaws}

@@

% This Plan Space Planning algorithm is translated into \eus
% representation as follows:

@@

\item[\textbf{Unbound Variable}]: a variable in the partial plan whose
  domain is not a singleton. Unbound Variables are resolved by value
  selection from their domain.

\item[\textbf{Open Condition}]: an open condition is an
  \texttt{INACTIVE} token. \texttt{INACTIVE} tokens can be generated
  by explicitly posted goals, or when a token is activated, rules may
  activate the creation of \texttt{INACTIVE} slave tokens as described
  in Section \ref{sec:europa:pr}. Open Conditions can be resolved in
  three ways:

@@

  \item[\textbf{Merging}]: A free activity is merged with a matching
    activity already in the plan. An activity $a$ is said to match an
    activity $a?$ iff $a$ and $a?$ unify \cite{russelnorvig} and the
    temporal constraints involving $a$ are satisfied by $a?$. Thus,
    $a$ and $a?$ can be considered the same activity without the need
    to introduce $a$ in the plan. Consequently, the domain rules associated with
    $a$ are not introduced into the planning process, because they have been already
    triggered when $a?$ was introduced in the plan.

  \item[\textbf{Activation}]: A new activity $a$ is introduced in the
    current plan associating it with the correct timeline, without
    choosing a specific time slot (if its timing causes a problem 
    it'll trigger a Threat flaw, explained below). 
    The domain rules associated with $a$ are applied and the subgoal
    activities resulting from those domain rules are introduced as inactive tokens. 
    This results in a number of open condition flaws, corresponding to
    the new subgoal activities.

  \item[\textbf{Rejection}]: Some goals may be optional; if so its
    corresponding flaw may be resolved by discarding the
    \texttt{INACTIVE} token.

@@

Algorithm \ref{alg:europa:solve} provides for a sound and complete
search, assuming that no flaws or available refinement operators are
pruned unnecessarily.  Dead-ends in the search are discovered through
constraint propagation, a vehicle for evaluating the consistency of a
partial plan and also for filtering infeasible values from
consideration prior to commitment. % This allows in some cases a strong
% look-ahead capability which is essential for tractable
% search.
Consistency testing is initiated by the \texttt{isInconsistent}
procedure. The algorithm permits a heuristically controlled search by
applying orderings for \texttt{chooseFlaw} and
\texttt{makeDecisionPoint} resulting in a chronologically-backtracked
search. Heuristics can be implemented by applying ordering/filtering
policies for \texttt{chooseFlaw} and \texttt{makeDecisionPoint}, for
instance, by dealing with earliest flaws first, or by making planning
decisions (open condition flaws) before scheduling decision (threats).

It should be emphasized that while this algorithm is commonly employed
with \eu implementations, it is only one of many that could be
implemented. As a simple example, consider the following solver for
the $N$-Queens problem, where an integer variable V$_i$ represents the
queen in column $i$; V$_1$ $= 3$ implies that the variable on column
$1$ is placed on row $3$.  Non-attack constraints are posted as
arithmetic constraints among the V$_i$ variables to ensure that all
the rows and diagonals contain only one queen. \eus Constraint
Reasoning Engine or \texttt{CRE} can then be used to compute the
non-attack constraint violations, with \texttt{ce.getViolations()} a
call on the API, to pick the queen that has he most violations.  By
implementing \texttt{getQueenWithMaxViolation()} and looking ahead to
see what moves are possible for a queen that would remove its
violations with the implementing of \texttt{getMoves()}, this solver
provides a viable planner implementation for a range of possible
initial states.

@@

>>>>

Fig. \ref{fig:europa-architecture} shows the main architectural
components in \eu and their relationships. The \textbf{Constraint
  Reasoning Engine} (\texttt{CRE}) manages domain types, variables and
constraints that define relationships among them. It also provides an
efficient arc consistency mechanism \cite{mackworth77}. The
\texttt{CRE} is designed so that specialized reasoning algorithms for
new constraints can be easily and efficiently plugged in. The
\textbf{Plan Database} (\texttt{PDB}) manages object and token
types. It also maintains the state of partial plans in terms of
object, variable and token instances. This constitutes the data
foundation on which end user applications and automated problem solvers
 are built. The \textbf{Rules Engine} (\texttt{RE})
modularly adds the semantics of domain rules that describe
dependencies between tokens. The \textbf{Solver Module} provides the
framework for flaw detection and resolution that is needed to
implement search algorithms for planning. The \textbf{API Layer}
provides access to services from all the different modules in a
programmatic way so that \eu can be easily embedded into specialized
applications.

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eue,
\texttt{NDDL} allows the user to specify the different components of a
problem domain in a precise and concise manner. Among the main
features that make \texttt{NDDL} a powerful tool for describing
problem domains are that it is object oriented. \texttt{NDDL} supports
classes and inheritance in similar fashion to popular Object-Oriented
languages like Java and C++. It also supports polymorphism for some of
the planning components. Using object classes and instances is a
time-tested approach to naturally describe problem domains
\cite{rumbaugh}. It offers declarative constructs to define
constraints and causality dependencies between actions and effects and
finally it offers procedural constructs to populate \eus \texttt{PDB}
with specific problem instances expressed in terms of objects,
variables, constraints, facts and goals. To articulate the most important
elements of \texttt{NDDL}, we show how the Shopping Agent problem
could be modeled:

@@

In the model above one can see how the main elements of a problem domain
can be mapped to \eus representation:

@@

\item[\textbf{Object Types}:] \texttt{NDDL} supports an object-oriented
  approach, where common data and behavior can be abstracted as a
  class. In this case, we have the \textit{Product} class represent
  the products that the Shopping Agent needs, \textit{Location}
  represents the places where products can be found,
  \textit{ProductLocation} represents the association arising when a
  product can be purchased at a location. The Shopping Agent itself is
  represented by an \textit{Agent} class, which relies on the
  \textit{AgentLocation} class to keep track of where the agent is at
  any point in time. \texttt{NDDL} supports a single rooted class
  hierarchy so that state and behavior can be extended through
  inheritance. We take advantage of that feature by making
  \textit{AgentLocation} inherit from \eus built-in \textit{Timeline}
  class which doesn't allow predicates to overlap in time for the same
  object instance. This feature is used for instance, to ensure that
  an agent can only be at one place at a time in our example.

\item[\textbf{Variable Types and Instances}:] All object attributes
  are represented as variables in this model. For instance, it uses
  integer variables to represent the built-in temporal parameters for
  each token (duration, start and end times), string
  variables to represent \textit{Location} and \textit{Product} names,
  and object reference variables to represent the structural
  relationships between different objects (for instance, each Agent
  maintains a variable to keep track of its \textit{AgentLocation}).

\item[\textbf{Predicate Types}:] Some classes like \textit{Product},
  \textit{Location} and \textit{ProductLocation} are static and so
  don't have any predicate or action types associated with them. The
  \textit{AgentLocation} class is used to keep track of state and so
  has \texttt{At} and \texttt{Going} predicate types associated with
  it. The \textit{Agent} class keeps track of the products it
  purchases through the \texttt{Own} predicate; having a separate
  \textit{AgentBag} class with its own predicates would have been an
  equally valid modeling choice.

\item[\textbf{Action Types}:] In our example, \textit{Agent} is solely
  the class that has behavior associated with it. Consequently, it
  defines action types to represent when the agent is taking action to
  \texttt{Go} to a \textit{Location}, or to \texttt{Buy} a
  \textit{Product}.

@@

\item[\textbf{Variables}:] A token type (predicate or action) is part
  of a class, therefore class attributes (see \texttt{object.location}
  above), token type parameters (see \texttt{[from, to]} in
  \texttt{Agent::Go}, or \texttt{[duration, product]} in
  \texttt{Agent::Buy}) and locally declared variables (see
  \texttt{possibleStores} in \texttt{Agent::Buy}) can all be used in
  the definition of a rule.

\item[\textbf{Constraints}:] Restrictions on values that variables can
  take within a rule are expressed through constraints. In the example
  above the \texttt{eq} constraint is used; in general, any constraint
  available in \eus library can be used.

\item[\textbf{Subgoals}:] In \texttt{Agent::Go}, subgoals are used to
  specify that the Shopping Agent's location must transition from
  origin to destination in a temporally consistent way. The
  \texttt{[meets, equals, metby]} operators (see Section
  \ref{sec:europa:inference}) are used to specify subgoals, along with
  the temporal relationship each token representing a subgoal has with
  the the token representing the \texttt{Agent::Go} action. Also,
  \texttt{[condition, effect]} annotations can be used when defining a
  subgoal inside an action. \eu does not intrinsicaly associate semantics
  with these annotations; but they can be extremely useful for search
  modules when creating plans.

@@

\end{verbatim}

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts are used to specify the current state of the
world, do not need any support and are represented as tokens in the
\texttt{PDB}. In this case we use facts to specify that the Shipping
Agent is initially at home. Goals are also represented as tokens in
the \texttt{PDB}. Different from facts, goals trigger inference in \eu
which determines whether flaws (unresolved subgoals, overloaded
resources, etc.) need to be resolved; if so, search modules can be
invoked to resolve those flaws as part of the deliberation
process. Finally, constraints are restrictions on and between facts 
and goals. In the example above, we use temporal constraints to
specify when facts hold and when the goals should be
accomplished. In the model above, \texttt{before} and \texttt{after} 
are \texttt{Allen Algebra} \cite{allen84} operators, which \eu supports to specify
temporal relations between tokens (specifically between their start and end timepoints)
as summarized in Fig. \ref{fig:allen-algebra}

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} \cite{allen84} relations shown
  above.}
\label{fig:allen-algebra}
\end{figure}

@@

In \eue, all variables and constraints from a problem domain
description are stored inside the \texttt{CRE}.  The \texttt{CRE}
provides a framework where any number of constraint types can
collaborate to perform inference and prune variable domains using
Bounds Propagation and Domain Reduction techniques \cite{marriott98}. 
\eu uses a Propagator pattern which keeps track of a set of
constraint instances of the same type and is responsible for ensuring
variables are updated through inference whenever a relevant change occurs
 (variable domain modifications or constraint addition/deletion).

In a planning problem, variable or constraint changes can come
directly from an end user (by introducing new activities or goals and
relating them to the rest of the plan through constraints, by changing
the start or end time of an activity, etc), or from an automated planner (by
expanding subgoals, introducing precedence constraints to resolve
mutual exclusion conflicts, etc). When these changes occur, each
propagator is only able to ensure consistency for the constraints that
are associated with it.  To maintain global consistency, the
\texttt{CRE} implements a variant of the \texttt{AC-3} algorithm
\cite{mackworth77} where propagators are invoked repeatedly until
the constraint network returns to a quiescent state.

@@

An \eu user can introduce new constraint types by implementing
specialized constraint and propagator classes; however, \eu provides a
built-in constraint library with a range of arithmetic, temporal, set
and resource constraint types so that a large number of problem
domains can be modeled without having to introduce new constraint
types. Tables \ref{tab:calconst}, \ref{tab:setconst},
\ref{tab:objhierarchy} and \ref{tab:misc} show the main constraint
types included in \eus Constraint Library.

@@

Temporal and resource constraints are especially important to model
real-world planning domains.

\eu uses variables to explicitly represent timepoints for plan
actions and states. Constraints among timepoints provide a natural
way to express domain axioms. For example, in order to state that
activity $A$ must occur before activity $B$ we can say that the end
timepoint of $A$ is $leq$ the start timepoint of $B$.
\cite{dechter91} proposed that constraints among timepoints can be
grouped together to form a Simple Temporal Network
(\texttt{STN}). Such a network can be transformed into a Distance
Graph (\texttt{DG}) where the outward arc from a node represents the
maximum distance from the source node to the target node.
Fig. \ref{fig:stn} illustrates an \texttt{STN} with $2$ variables and
a single constraint along with the resulting \texttt{DG}.
\cite{dechter91} also showed that shortest path algorithms could be
used to propagate values in the network and discover a negative cycle:
namely a path from a node to itself that has a path length less than
$0$. If such a cycle exists, the network is inconsistent. It was
further shown that a single-source shortest path algorithm was
sufficient to detect a negative cycle and provide sufficient
propagation to yield a backtrack-free search. \eus propagator for
temporal constraints is a straight implementation of this algorithm.

@@

Most real-world planning problems need to deal with plans made up of
actions that typically depend upon limited resources (\eg fuel,
storage capacity, battery state of charge, data buffers).  \eu
provides built-in support for resource computation. In \eus
representation resource object types come in three forms with related
constraints. \textbf{Reservoir} resources can have both
\texttt{consume(resource,time,quantity)} and
\texttt{produce(resource,time,quantity)} constraints; production and
consumption is instantaneous. \textbf{Reusable} resources on the other
hand, have constraints that use a quantity of the resource for a
specific duration (\texttt{uses(resource,start,duration,quantity)}),
which are then subsequently released. \textbf {Unary} resources are
reusables with capacity$ = 1$, which allows \eu to perform run time
savings in comparison to the generic reusable class. Configuration
options are offered such that in addition to the usual constraint
violations when a resource is completely depleted, plan flaws can be
generated whenever a resource level is outside a specific upper or
lower bound, or when instantaneous production or consumption exceeds a
specific threshold.

Since \eu supports a flexible temporal representation, where time is
represented by intervals, instead of single values, determining
whether a resource is the source of a plan flaw or a constraint
violation in general is substantially more difficult than just
computing a profile of the resource and checking against some limits.
With temporal flexibility, a number of resource profiles could be
possible depending on precisely when each production/consumption event
occurs out of a range of possible values that are in the domain of the
variables that represent the temporal extent for the event.
\cite{Muscettola04,Muscettola06} proposed an algorithm where the
variables that represent the time for production/consumption events
are placed in a temporal network, similar to the network used by the
temporal propagator; constraints on the resource can then be seen as
quantities flowing through that network. Computing the maximum flow
then, yields the maximum resource usage and thus a lower bound on the
resource level; computing a minimum flow will similarly yield an upper
bound. Since maximum flow on a temporal network can be computed in
polynomial time, \eu takes advantage of this algorithm to efficiently
compute upper and lower bound envelopes on the resource
level. Fig. \ref{fig:resenvelopes} shows an example of such a resource
computation.

@@

Resource envelopes can be useful when trying to evaluate plan
brittleness in the face of changing activity durations or start
times. They can also be useful when searching for plans; however they
are relatively expensive to compute. Therefore their ability to prune
the search space must be weighted against the ability to search faster
\cite{6007772}, \comment{cite "Stable Grounded Inference in Flexible Resource Scheduling" by Morris, Bresina, Barreiro}.

@@

Given the above, an automated problem solver can be created using a
\emph{domain model} that describes the variable, object, predicate and
action types that are relevant for the problem, a \emph{problem
  instance} or initial state that consists of variable and object
instances that exist for the entire planning horizon, \emph{temporally
  scoped predicate and action} instances and \emph{temporally scoped
  goals}. This information is retained in \eus \texttt{PDB} so that
inference and search mechanisms can be used to look for a problem
solution.

Since temporal intervals may extend indefinitely in both positive and
negative directions, for problems with a temporal dimension it is also
common to specify a \emph{planning horizon} when invoking a solver;
any decisions that fall completely outside of the horizon can be
ignored. This is useful to ignore activities that may occur far in the future 
or in the past to be relevant for a particular plan, ensuring that the solver is making
as few commitments as necessary and reducing the complexity of plan generation.
The \emph{planning horizon} is also useful when dealing with recurrent activities;
consider for instance a model where a medical checkup has to occur
every $3$ months. The user can model a recurring activity and then use
the horizon to ensure that a finite number of medical checkup
activities are generated as part of a plan. 
\subsubsection{Flaw Resolution}
\label{sec:europa:flaws}

@@

@@

\item[\textbf{Unbound Variable}:] a variable in the partial plan whose
  domain is not a singleton. Unbound Variables are resolved by value
  selection from their domain.

\item[\textbf{Open Condition}:] an open condition is an
  \texttt{INACTIVE} token. \texttt{INACTIVE} tokens can be generated
  by explicitly posted goals, or when a token is activated, rules may
  cause the creation of \texttt{INACTIVE} slave tokens as described
  in Section \ref{sec:europa:pr}. Open Conditions can be resolved in
  three ways:

@@

  \item[\textbf{Merging}:] An \texttt{INACTIVE} token is merged with a
    matching \texttt{ACTIVE} token already in the plan. A token $a$ is said to
    match another token $a?$ iff $a$ and $a?$ unify \cite{russelnorvig}
    and the temporal constraints involving $a$ are satisfied by
    $a?$. Thus, $a$ and $a?$ can be considered the same token
    without the need to introduce $a$ in the plan. Consequently, the
    domain rules associated with $a$ are not introduced into the
    planning process, since they are already triggered when $a?$ was
    introduced in the plan.

  \item[\textbf{Activation}:] A token $a$ is introduced in the
    current plan associating it with the appropriate timeline, without
    choosing a specific time slot; if its timing causes a problem a
    Threat flaw will be triggered (see below).  The domain rules
    associated with $a$ are applied and the subgoal activities
    resulting from those domain rules are introduced as \texttt{INACTIVE}
    tokens.  This results in a number of open condition flaws,
    corresponding to the new subgoal activities.

  \item[\textbf{Rejection}:] Some goals may be optional; if so its
    corresponding flaw may be resolved by discarding the
    \texttt{INACTIVE} token.

@@

Algorithm \ref{alg:europa:solve} provides for a sound and complete
search, assuming that no flaws or available refinement operators are
pruned unnecessarily.  Dead-ends in the search are discovered through
constraint propagation, a vehicle for evaluating the consistency of a
partial plan and also for filtering infeasible values from
consideration prior to commitment. Consistency testing is initiated by the \texttt{isInconsistent}
procedure.

The algorithm permits a heuristically controlled search by
applying orderings for \texttt{chooseFlaw} and
\texttt{makeDecisionPoint} resulting in a chronologically-backtracked
search. Heuristics can be implemented by applying ordering/filtering
policies for \texttt{chooseFlaw} and \texttt{makeDecisionPoint}, for
instance, by dealing with earliest flaws first, or by making planning
decisions (open condition flaws) before scheduling decision (threats).

It should be emphasized that while this algorithm is commonly employed
by applications that embed  \eu, it is only one of many that could be
implemented. As a simple example, consider the following solver for
the $N$-Queens problem, where an integer variable V$_i$ represents the
queen in column $i$; V$_1$ $= 3$ implies that the variable on column
$1$ is placed on row $3$.  Non-attack constraints are posted as
arithmetic constraints among the V$_i$ variables to ensure that all
the rows and diagonals contain only one queen.

@@

\eus Constraint Reasoning Engine or \texttt{CRE} can be used to compute the
non-attack constraint violations, with \texttt{ce.getViolations()} a
call on the \eu API. The \texttt{CRE} can also be used to pick the queen 
that has he most violations in \texttt{getQueenWithMaxViolation()}, and to
look ahead and see what moves are possible for a queen that would remove its
violations in \texttt{getMoves()}.
